Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-12-13 manuscript/manuscript.tex
Improve delta ratio determination for exp density

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/exponential-delta.pdf}
\caption{
    Numerical error for different exponential density functions and values of the delta
    ratio.
    The top panels show the exponential density functions for different values of $b$
    (Eq.~\ref{eq:density-exp}).
    The error computations were performed on the ``Big Grid'' configuration described in
    Table~\ref{tab:grids}, with fixed values of the distance-size ratio
    $D$ obtained for the linear density case ($D=1, 2, 8$ for the
    potential, gradient components, and Marussi tensor components, respectively).
    The difference is reported as a percentage of the shell values.
    }
\label{fig:delta-exponential}
\end{figure*}

Having chosen values of $D$ equal to the ones obtained for the linear density case, we
are free to explore the integration error as a function of $\delta$ in more detail and
how it varies for different values of $b$ (Eq.~\ref{eq:density-exp}).
Now we perform the error calculations for the every grid detailed in
Table~\ref{tab:grids} for the 20 spherical shells models detailed in
Table~\ref{tab:tess-collections}.
The calculations are repeated for different values of $b$ to examine the error variation
with the sharpness of the density function.
Because larger $\delta$ values result in fewer tesseroid divisions,
our intention is to find the highest value of $\delta$ whose numerical error is bellow
the $0.1\%$ threshold.

Fig.~\ref{fig:delta-exponential} shows the density functions and the resulting relative
differences ($V$, $g_z$, and $g_{zz}$) for different values of $b$ and thickness of the
shell and tesseroid model.
The relative difference for $V$ and $g_z$ falls below the 0.1\% threshold for
$\delta = 0.2$ in both the thick and thin shell cases.
On the other hand, optimal values of $\delta$ for $g_{zz}$
are $\delta = 0.2$ for the thin shell and $\delta = 0.01$ for the thick shell.
Once again, we opt for the conservative value of $\delta = 0.01$ and sacrifice
performance for the sake of accuracy.

>>>>

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/exponential-density-diffs.pdf}
\caption{
    Numerical error for different exponential density functions and values of the delta
    ratio.
    The percentage difference values were obtained from the comparison between the
    analytical solution and the numerical approximation of the gravitational fields ($V$
    and $g_z$) generated by a spherical shell with an exponential density function and
    different values of $b$ (Eq.~\ref{eq:density-exp}).
    These comparisons were carried out on every grid defined on Table~\ref{tab:grids},
    making use of the 20 shell models detailed on Table~\ref{tab:tess-collections} and
    fixing the distance-size ratio $D$ to the values obtained for the linear density
    case ($D = 1, 2.5$ for the potential and $g_z$, respectively).
    The difference is reported as a percentage of the shell values.
    To improve visualization of the results, each curve corresponds to the maximum
    difference between every shell model.
    }
\label{fig:delta-exponential}
\end{figure*}

Having chosen values of $D$ equal to the ones obtained for the linear density case, we
are free to explore the integration error as a function of $\delta$ in more detail and
how it varies for different values of $b$ (Eq.~\ref{eq:density-exp}).
Now we perform the error calculations for the every grid detailed in
Table~\ref{tab:grids} and for the 20 spherical shells models detailed in
Table~\ref{tab:tess-collections}.
The calculations are repeated for different values of $b$ (1, 2, 5, 10, 30, 100) to
examine the error variation with the sharpness of the density function.
Because larger $\delta$ values result in fewer tesseroid divisions,
our intention is to find the highest value of $\delta$ whose numerical error is bellow
the $0.1\%$ threshold.

Fig.~\ref{fig:exp-densities} shows the density functions for each value of $b$ and
Fig.~\ref{fig:delta-exponential} shows the resulting relative
differences ($V$ and $g_z$).
In order to improve the visualization of the results, each curve corresponds to the
maximum difference between all shell models.
The relative difference for $V$ and $g_z$ falls below the 0.1\% threshold
for each computation grid, shell model and $b$ value if $\delta = 0.1$.
These results also show that there is no significant relation between the $b$
value and the numerical error, proving that the density-based discretization algorithm
is capable of limiting the error due to the density variation regardless the variation
of the exponential density.
